Think deep for reef resilience

### 2025-08-17 paper/paper_draft.qmd
Update carbonate budget results to reflect joint figure

**\[TI: Ben, Just noticed - we need to make the rows (Management areas) go from North to
South!\]**

@@

Reef median depth has a moderate positive correlation with the number of years a reef
maintains coral cover above the assumed carbonate budget threshold required to sustain live
coral cover ($\theta_{0}$). Greater median depth is a key characteristic defining reefs that
sustain coral cover above $\theta_{0}$ [@fig-access-esm1-5-depth-carbonate]. Unsurprisingly,
the number of years above $\theta_0$ substantially decreases as $\theta_0$ is increased, as
does the Spearman's correlation. This is likely due to an increasing number of reefs being
unable to maintain coral cover above $\theta_0$ as it is increased, with the distribution of
years above threshold becoming increasingly right-skewed
[@fig-access-esm1-5-depth-carbonate; @fig-access-esm1-5-conn-carbonate]. In contrast,
log10-transformed total connectivity strength ($\text{log10}(S)$) was found to have no
correlation with the number of years a reef maintains cover above the required thresholds
[@fig-access-esm1-5-conn-carbonate]. The correlation between $\text{log10}(S)$ and
$\theta_0$ slightly decreases as the assumed threshold for $\theta_0$ increases.

![Number of years reef coral cover exceeds the carbonate budget threshold from 2025 to 2075.
The assessed carbonate budget thresholds range from 10 to 20% of total reef area. Bracketed
x-axis values represent the Spearman's correlation between reef depth and the number of
years cover exceeds the carbonate budget threshold. Point colour represents the median reef
depth in metres. Red points represent the mean number of years above an assumed carbonate
budget threshold across reefs. Note that deeper reefs are better able to maintain cover
above the indicative carbonate budget threshold. Displays results from median time series
using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate
fig-align="center" width="14.82cm"}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2075. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef base-10 log of
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. The colour of each point represents the base-10 log of scaled reef total
connectivity strength. Red points represent the mean number of years above a carbonate
threshold across reefs. Displays results from median time series using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#fig-access-esm1-5-conn-carbonate
fig-align="center" width="14.82cm"}

@@

The results presented suggest that when modelled as part of a large and complex ecosystem,
reef depth is a key factor contributing to ongoing reef resilience. We provide evidence in
support of depth as a form of thermal refugia for shallow-water reefs on the GBR. While
protecting ‘key node’ reefs in larval connectivity networks is an important consideration
for reef restoration efforts [@anthony2020; @hock2017], current connectivity modelling
methods, and our modelling framework, suggest that connectivity may not be as important for
reef resilience under projected conditions as factors such as depth **\[TI: TO BE
REVISITED\]**. We suggest further analyses that investigate the benefits of incorporating
depth-based resilience into reef intervention decision making, as well as further research
and development in connectivity modelling methods for whole-ecosystem scales.

>>>>

@@

Reef median depth has a moderate positive correlation with the number of years a reef
maintains coral cover above the assumed carbonate budget threshold required to sustain live
coral cover ($\theta_{0}$). Greater median depth is a key characteristic defining reefs that
sustain coral cover above $\theta_{0}$ [@fig-access-esm1-5-depth-carbonate]. Unsurprisingly,
the number of years above $\theta_0$ substantially decreases as $\theta_0$ is increased, as
does the Spearman's correlation. This is likely due to an increasing number of reefs being
unable to maintain coral cover above $\theta_0$ as it is increased, with the distribution of
years above threshold becoming increasingly right-skewed
[@fig-access-esm1-5-depth-carbonate]. In contrast, log10-transformed total connectivity
strength ($\text{log10}(S)$) was found to have no correlation with the number of years a
reef maintains cover above the required thresholds
[@fig-access-esm1-5-depth-conn-carbonate]. The correlation between $\text{log10}(S)$ and
$\theta_0$ slightly decreases as the assumed threshold for $\theta_0$ increases.

![Number of years reef coral cover exceeds the carbonate budget threshold from 2025 to 2075.
The assessed carbonate budget thresholds range from 10 to 20% of total reef area. Bracketed
x-axis values represent the Spearman's correlation between reef depth and the number of
years cover exceeds the carbonate budget threshold. Point colour represents the median reef
depth in metres in the left-hand panel, and log total connectivity strength in the
right-hand panel. Red points represent the mean number of years above an assumed carbonate
budget threshold across reefs. Deeper reefs are better able to maintain cover above the
indicative carbonate budget threshold, while there is no considerable relationship between
connectivity strength and carbonate budget maintenance. Displays results from median time
series using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/depth_conn_carbonate_comparison.png){#fig-access-esm1-5-depth-conn-carbonate
fig-align="center" width="14.82cm"}

@@

The results presented suggest that when modelled as part of a large and complex ecosystem,
reef depth is a key factor contributing to ongoing reef resilience. We provide evidence in
support of depth as a form of thermal refugia for shallow-water reefs on the GBR. While
protecting ‘key node’ reefs in larval connectivity networks is an important consideration
for reef restoration efforts [@anthony2020; @hock2017], current connectivity modelling
methods, in combination with our modelling framework, suggest that connectivity may not be
as important for reef resilience under projected conditions as environmental factors such as
depth and thermal stress levels. We suggest further analyses that investigate the benefits
of incorporating depth-based resilience into reef intervention decision making, as well as
further research and development in connectivity modelling methods for whole-ecosystem
scales.
